working title Compatibility system and stigma size are the main predictors of heterospecific pollen effect

### 2019-02-21 Manuscript_draft/Manuscript_draft.Rmd
more changes in intro

Plant traits are crucial to understand heterospecific pollen effect but the multifactorial nature of the traits that are involve in the pollen-pistil interaction make difficult to unravel what are the main traits in driving the effect. On the one hand the main predictive traits of effect from the donor perspective are pollen size, pollen allelopathy and pollen apertures (REFS). In Ashamn 2013 small pollen is predicted to cause a greater fitness decrease, although this can be true there are also other possibilities to consider which can obscure a predictive framework like big pollen can clogg small stigmas with fewer pollen grains, bigger stigmas are less likely to be clogged by small pollen grains and bigger pollen can outcompete smaller pollen grains due to faster pollen tube growth rate (REF). On the other hand  the main recipient traits that can drive heterospecific pollen effect are: stigma size, style length, number of ovules, incompatibility system and flower morphology (REFS). In first place, greater stigmatic area is related with greater amount of heterospecific pollen (REF) and therefore possibly a greater negative effect (REF). Species with shorter styles are more likely to be vulnerable to heterospecific pollen effect (REF). For species that are self-incompatible have stronger barriers to heterospecific pollen than self-compatible species @ashman2013. Nonetheless, an effect of foreign pollen is a bit obscured by the variability within species, however species that are strong selfers or strong outcrossers have less variablity in mating systems and predictions of effect could be more realistic (see figure 1 from @whitehead2018). All in all, there is a need to deepen in the understanding of the relative importance of these traits, due to we still do not know what are the main traits in driving the effect with all these possible combinations that can occur in nature.

[Comments on it](Romina)

Species closely related are more likely to have similar traits XXXXXXXXX.  (Refs? Brown and Mitchell (2001) Arceo-Gómez et al. (2016) Tong and Huang (2016) ).  Several studies predict that the impact of HP transfer is likely to be greater for closely related species (Ashman and Arceo-Gómez (2013)).  Few studies however, have focused on the impacts of heterospecific pollen of distantly related species Thomson et al. (1982) Galen and Gregory (1989) Neiland and Wilcock (1999). Yet, most insects and most stigmas have been found to carry multiple species of foreign pollen with little attention to degree of relatedness (Arceo-Gómez and Ashman (2016); Fang and Huang (2013) ; also cite studies from pollen transfer networks here such as... ).

Further, a majority of plant species are generalist and thus receive visits from multiple different pollinators.  Given these are generally the ones that receive greater loads of heterospecific pollen Fang and Huang (2013) and unrelated species are more likely to coexist  with other species due to less niche overlap (Ref),  understanding the role of foreign pollen from distantly related species thus deserves greater attention in understanding coexistence blah blahXXXXX refs.. Notwithstanding, the effect of heterospecific pollen of far and close related species at community level remains to be explored beyond single pairwise interactions. In this study we investigated how floral reproductive traits and relatedness mediate the impact of HP transfer by asking the following research questions : To what extent do (i) floral reproductive traits and (ii) relatedness, mediate the impacts of heterospecific pollen on seed set.  We do this by creating an artificial co-flowering community with 10 species belonging to three different families with different traits.

@@

>>>>

Plant traits are crucial to understand heterospecific pollen effect but the multifactorial nature of the traits that are involve in the pollen-pistil interaction make difficult to unravel what are the main traits in driving the effect. These traits can be seen from a male perspective of both donor and recipient where pollen size, pollen aperture number and pollen allelopathy are key components to understand the outcome of foreign pollen arrival (REFS). In @ashman2013 small pollen is predicted to cause a greater fitness decrease, although this can be true there are also other possibilities to consider which can obscure a predictive framework like big pollen can clogg small stigmas with fewer pollen grains, bigger stigmas are less likely to be clogged by small pollen grains and bigger pollen can outcompete smaller pollen grains due to faster pollen tube growth rate @williams1990. Moreover, to understand the different mechanical or chemical effects of pollen also the female traits of the pollen recipient have to be considered, from the literature these main traits are: stigma size, style length, number of ovules, incompatibility system and finally a structural trait such as flower morphology @montgomery2012; @ashman2013; @tong2016. For example, greater stigmatic area is positively correlated with greater heterospecific pollen deposition @montgomery2012 and therefore possibly with an increase in negative effect. For species that are self-incompatible the barriers towards heterospecific pollen are stronger than self-compatible species @ashman2013. Nonetheless, an effect of foreign pollen is a bit obscured by the variability within species, however species that are strong selfers or strong outcrossers have less variablity in mating systems and predictions of effect could be more realistic (see figure 1 from @whitehead2018). Although past research has progress in the understanding of what traits can mediate the effect as we have shown here, there are multiple traits involved and multiple possible scenarios still to be explored empirically for a full understanding of the importance of heterospecific pollen effect in nature.

Closely related species are more likely to have similar traits (Refs? Brown and Mitchell (2001) Arceo-Gómez et al. (2016) Tong and Huang (2016)). Several studies predict that the impact of HP transfer is likely to be greater for closely related species (Ashman and Arceo-Gómez (2013)).  Few studies however, have focused on the impacts of heterospecific pollen of distantly related species Thomson et al. (1982) Galen and Gregory (1989) Neiland and Wilcock (1999). Yet, most insects and most stigmas have been found to carry multiple species of foreign pollen with little attention to degree of relatedness (Arceo-Gómez and Ashman (2016); Fang and Huang (2013); also cite studies from pollen transfer networks here such as... ). Traditionally, studies have overlook the effect of closely related species but in a natural context both can occur and different degree of relatedness should be considered.

Further, a majority of plant species are generalist and thus receive visits from multiple different pollinators.  Given these are generally the ones that receive greater loads of heterospecific pollen Fang and Huang (2013) and unrelated species are more likely to coexist  with other species due to less niche overlap (Ref),  understanding the role of foreign pollen from distantly related species thus deserves greater attention in understanding coexistence blah blahXXXXX refs.. Notwithstanding, the effect of heterospecific pollen of far and close related species at community level remains to be explored beyond single pairwise interactions. Moreover, heterospecific pollen studies in nature have the complexity added great environmental variability which can lead to confounding interpretations and diverse floral structures and some of them complex such as Asteraceae species, which involve tedious work to be able to study foreign pollen effect. For this reason, we investigated how floral reproductive traits and relatedness mediate the impact of heterospecific transfer by asking the following research questions : To what extent do (i) floral reproductive traits and (ii) relatedness, mediate the impacts of heterospecific pollen on seed set.  We do this by creating an artificial co-flowering community with 10 species belonging to three different families with different traits.

-Would like to add here floral complexity

-Environmental variability of studies on the field

@@

Herbs vs tress, annual vs perennial... Many flowers vs few flowered species; structural composition on a system

### 2019-02-21 Manuscript_draft/Manuscript_draft.tex
more changes in intro

Plant traits are crucial to understand heterospecific pollen effect but
the multifactorial nature of the traits that are involve in the
pollen-pistil interaction make difficult to unravel what are the main
traits in driving the effect. On the one hand the main predictive traits
of effect from the donor perspective are pollen size, pollen allelopathy
and pollen apertures (REFS). In Ashamn 2013 small pollen is predicted to
cause a greater fitness decrease, although this can be true there are
also other possibilities to consider which can obscure a predictive
framework like big pollen can clogg small stigmas with fewer pollen
grains, bigger stigmas are less likely to be clogged by small pollen
grains and bigger pollen can outcompete smaller pollen grains due to
faster pollen tube growth rate (REF). On the other hand the main
recipient traits that can drive heterospecific pollen effect are: stigma
size, style length, number of ovules, incompatibility system and flower
morphology (REFS). In first place, greater stigmatic area is related
with greater amount of heterospecific pollen (REF) and therefore
possibly a greater negative effect (REF). Species with shorter styles
are more likely to be vulnerable to heterospecific pollen effect (REF).
For species that are self-incompatible have stronger barriers to
heterospecific pollen than self-compatible species Ashman and
Arceo-Gómez (2013). Nonetheless, an effect of foreign pollen is a bit
obscured by the variability within species, however species that are
strong selfers or strong outcrossers have less variablity in mating
systems and predictions of effect could be more realistic (see figure 1
from Whitehead et al. (2018)). All in all, there is a need to deepen in
the understanding of the relative importance of these traits, due to we
still do not know what are the main traits in driving the effect with
all these possible combinations that can occur in nature.

\href{Romina}{Comments on it}

Species closely related are more likely to have similar traits
XXXXXXXXX. (Refs? Brown and Mitchell (2001) Arceo-Gómez et al. (2016)
Tong and Huang (2016) ). Several studies predict that the impact of HP
transfer is likely to be greater for closely related species (Ashman and
Arceo-Gómez (2013)). Few studies however, have focused on the impacts of
heterospecific pollen of distantly related species Thomson et al. (1982)
Galen and Gregory (1989) Neiland and Wilcock (1999). Yet, most insects
and most stigmas have been found to carry multiple species of foreign
pollen with little attention to degree of relatedness (Arceo-Gómez and
Ashman (2016); Fang and Huang (2013) ; also cite studies from pollen
transfer networks here such as\ldots{} ).

Further, a majority of plant species are generalist and thus receive
visits from multiple different pollinators. Given these are generally
the ones that receive greater loads of heterospecific pollen Fang and
Huang (2013) and unrelated species are more likely to coexist with other
species due to less niche overlap (Ref), understanding the role of
foreign pollen from distantly related species thus deserves greater
attention in understanding coexistence blah blahXXXXX refs..
Notwithstanding, the effect of heterospecific pollen of far and close
related species at community level remains to be explored beyond single
pairwise interactions. In this study we investigated how floral
reproductive traits and relatedness mediate the impact of HP transfer by
asking the following research questions : To what extent do (i) floral
reproductive traits and (ii) relatedness, mediate the impacts of
heterospecific pollen on seed set. We do this by creating an artificial
co-flowering community with 10 species belonging to three different
families with different traits.

@@

@@

>>>>

Plant traits are crucial to understand heterospecific pollen effect but
the multifactorial nature of the traits that are involve in the
pollen-pistil interaction make difficult to unravel what are the main
traits in driving the effect. These traits can be seen from a male
perspective of both donor and recipient where pollen size, pollen
aperture number and pollen allelopathy are key components to understand
the outcome of foreign pollen arrival (REFS). In Ashman and Arceo-Gómez
(2013) small pollen is predicted to cause a greater fitness decrease,
although this can be true there are also other possibilities to consider
which can obscure a predictive framework like big pollen can clogg small
stigmas with fewer pollen grains, bigger stigmas are less likely to be
clogged by small pollen grains and bigger pollen can outcompete smaller
pollen grains due to faster pollen tube growth rate Williams and Rouse
(1990). Moreover, to understand the different mechanical or chemical
effects of pollen also the female traits of the pollen recipient have to
be considered, from the literature these main traits are: stigma size,
style length, number of ovules, incompatibility system and finally a
structural trait such as flower morphology Montgomery and Rathcke
(2012); Ashman and Arceo-Gómez (2013); Tong and Huang (2016). For
example, greater stigmatic area is positively correlated with greater
heterospecific pollen deposition Montgomery and Rathcke (2012) and
therefore possibly with an increase in negative effect. For species that
are self-incompatible the barriers towards heterospecific pollen are
stronger than self-compatible species Ashman and Arceo-Gómez (2013).
Nonetheless, an effect of foreign pollen is a bit obscured by the
variability within species, however species that are strong selfers or
strong outcrossers have less variablity in mating systems and
predictions of effect could be more realistic (see figure 1 from
Whitehead et al. (2018)). Although past research has progress in the
understanding of what traits can mediate the effect as we have shown
here, there are multiple traits involved and multiple possible scenarios
still to be explored empirically for a full understanding of the
importance of heterospecific pollen effect in nature.

Closely related species are more likely to have similar traits (Refs?
Brown and Mitchell (2001) Arceo-Gómez et al. (2016) Tong and Huang
(2016)). Several studies predict that the impact of HP transfer is
likely to be greater for closely related species (Ashman and Arceo-Gómez
(2013)). Few studies however, have focused on the impacts of
heterospecific pollen of distantly related species Thomson et al. (1982)
Galen and Gregory (1989) Neiland and Wilcock (1999). Yet, most insects
and most stigmas have been found to carry multiple species of foreign
pollen with little attention to degree of relatedness (Arceo-Gómez and
Ashman (2016); Fang and Huang (2013); also cite studies from pollen
transfer networks here such as\ldots{} ). Traditionally, studies have
overlook the effect of closely related species but in a natural context
both can occur and different degree of relatedness should be considered.

Further, a majority of plant species are generalist and thus receive
visits from multiple different pollinators. Given these are generally
the ones that receive greater loads of heterospecific pollen Fang and
Huang (2013) and unrelated species are more likely to coexist with other
species due to less niche overlap (Ref), understanding the role of
foreign pollen from distantly related species thus deserves greater
attention in understanding coexistence blah blahXXXXX refs..
Notwithstanding, the effect of heterospecific pollen of far and close
related species at community level remains to be explored beyond single
pairwise interactions. Moreover, heterospecific pollen studies in nature
have the complexity added great environmental variability which can lead
to confounding interpretations and diverse floral structures and some of
them complex such as Asteraceae species, which involve tedious work to
be able to study foreign pollen effect. For this reason, we investigated
how floral reproductive traits and relatedness mediate the impact of
heterospecific transfer by asking the following research questions : To
what extent do (i) floral reproductive traits and (ii) relatedness,
mediate the impacts of heterospecific pollen on seed set. We do this by
creating an artificial co-flowering community with 10 species belonging
to three different families with different traits.

-Would like to add here floral complexity

-Environmental variability of studies on the field

@@

Herbs vs tress, annual vs perennial\ldots{} Many flowers vs few flowered
species; structural composition on a system

@@

\hypertarget{ref-williams1990}{}
Williams, E., and J. Rouse. 1990. Relationships of pollen size, pistil
length and pollen tube growth rates in rhododendron and their influence
on hybridization. Sexual Plant Reproduction 3:7--17.
